Transmembrane helices are also an overlooked source of major histocompatibility complex class II epitopes and evolutionary more conserved than expected by chance

### 2020-04-29 article.tex
HLA classification and epitope prediction

\richel{Check for correctness}
Our immune system fights invaders on a daily basis.
These invaders can be fungi, bacteria or viruses.
The innate immune response is its first general 
and immediate strategy, where the acquired immune response
needs time to develop its specialized and more effective
combat forces.

@@

\richel{Check for correctness}
All nucleated cells in humans present randomly sampled polypeptides
fragments to the surroundings of the cell using MHC class I molecules.
If a cell gets infected by a virus, also the virus' polypeptides
will be presented at the cell surface. The viral antigen is detected 
by cytotoxic T lymphocytes, which will kill the infected cell.

@@

\richel{Check for correctness}
Bacteria and fungi do not infect a cell with their foreign DNA,
but are detected by the proteins on their cell walls. MHC class II molecules 
bind to specific polypeptide fragments. This binding allows T and B cells,
which all express MHC-II, to detect a pathogen and trigger the response to
kill the pathogen. Note that the foreign proteins of a viral envelope 
can be detected in this way as well, which is commonly used in vaccination.

\paragraph{MHCs bind hydrophobic epitopes}

\richel{Check for correctness}
Any human's immune system detects only a fraction of all possible
polypeptide fragments. 
Already each MHC molecule can only bind a
subset of all possible polypeptides. 
The polypeptides that can bind are also likelier to by hydrophobic,
as most TMHs have a hydrophobic binding pocket \richel{Reference needed}.
There are two sources of hydrophobic polypeptides, which are 
from membrane proteins and of soluble proteins.
Membrane proteins have at least one TMH to anchor them to
the membrane. TMHs are hydrophobic, as this is required to span the 
hydrophobic cellular lipid membrane. Additionally,
they often have a length of 23 amino acids to be able to span
the membrane.
Soluble proteins have (non-TMH) hydrophobic regions as well, 
to guides the protein to achieve its 3D configuration.

@@

\richel{Check for correctness}
For MHC-I, it was found that predicted epitopes derived 
from human transmembrane helices (TMHs)
are over-presented by all HLA-A and 
most HLA-B super types (\cite{bianchi2017}). 
One explanation is that the presentation of TMHs 
may have an evolutionary advantage for 
the (human) host, as TMHs have a reduced variability 
due to the functional requirement of being able to span a lipid 
bilayer. 
Due to this, pathogens have a lower chance to develop an escape mutation,
as many mutations will result in a disfunctional TMH.
Note that the mechanism by which a cell presents its TMHs is
yet unknown.

@@

\richel{Check for correctness}
If presentation of TMHs on MHC-I would bring an evolutionary advantage 
in the recognition of pathogens by the immune system, 
it would follow that this is equally important for MHC-II, 
especially as the help of CD4+ T cells is needed for a long lasting CD8+ T cell 
response \cite{novy2007cd4}. 
The mechanism to detect foreign TMHs by MHC-II would be unknown, 
similar to the discovery that MHC-I presents TMHs.

@@

\richel{Check for correctness}
The human immune system and human pathogen are in an evolutionary
arms race: our immune systems is selected for the detection
of pathogens, whereas pathogens are selected to avoid detection.
From a pathogen's point of view, however, this struggle 
is of only minor importance:
in seasonal influenza, for example, the selection pressure
exerted by the immune system was only limited \cite{han2019individual}
\richel{but this study does not look into the evolution of TMHs}.
In general, on would expect that evolutionary selection is strongest 
for regions that are important for viral proteins' functions:
in COVID-19, for example, there is preliminary evidence that the strongest
selection pressure is upon residues that changes its virulence \cite{velazquez2020positive}
\richel{residue 251 of ORF3a (an ion channel and a 
viroporin, residue is not part of TMH), and position 84 of ORF8 (unknown function)}.
The part of a pathogen's proteome that is essential, 
only accounts for small part of its full proteome.
Also, these essential parts differ widely between pathogens.
Because of this scarcity and variance in targets, 
one expects that the human immune system
is not tailored to detect these sites, 
as hinted by upon by the aforementioned influenza study.

@@

\richel{Check for correctness}
Regarding human viruses, we do not know the selection pressure
exerted by the human immune system.
For viral TMHs that are undetected by most human haplotypes, we
expect these to be conserved, to remain avoiding detection.
For viral TMHs that are detected by most human haplotypes, we
expect there to be a higher mutation rate, to avoid detection.
Additionally, viruses are not only selected for their evasion of an 
immune response, yet are selected to have a high reproductive value.
Because the strength of the evolutionary signal is unknown,
we use a strong data set to be able to detect it.

@@

\richel{Check for correctness}
Regarding human pathogenic bacteria, we do not know the extent to which the 
human immune systems selects on these, nor which loci are selected
upon \richel{so I should read the literature on that}.
Because most bacterial pathogens are generalists, that is,
can infect multiple hosts, we expect the selection pressure exerted
by the human immune response to be weak.
For bacterial TMHs that are undetected by most human haplotypes, we
expect these to be conserved, to remain avoiding detection.
For bacterial TMHs that are detected by most human haplotypes, we
expect there to be a higher mutation rate, to avoid detection.
Also, it is unknown if bacterial TMHs are detected by MHC-II at all.
Due to this, the strength of the evolutionary signal is unknown,
thus we need a strong data set to be able to detect it.

@@

\richel{Check for correctness}
To detect an evolutionary signal for selection on the TMHs of viruses, 
we use the many time-dated COVID-19 proteomes.
We can assume that only after the moment that COVID-19 spilled over 
to humans, it has mostly been the human immune system selecting against
its detection, although this is only part of the selection for 
having a high transmission rate.
To detect an evolutionary signal for selection on the TMHs of bacteria, 
we use the proteomes of Mycobacterium. Mycobacterium is \richel{something}.

@@

\richel{Check for correctness}
For MHC-II, we do the same analysis as MHC-I, except for
MHC-II haplotypes. We picked the top 21 MHC-II alleles that are most abundant 
in the current human population (\cite{greenbaum2011functional}).
\frans{Show coverage}
We chose to use 21 alleles to have a similar number of alleles as is done
for MHC-I \richel{Unsure if this was the actual reason}.
Same as \cite{bianchi2017}, we defined the 5\% of polypeptides 
with the lowest IC50 values as binders. We then counted the percentage
of binders that have one or more amino acids that are part of a TMH.

@@

>>>>

Our immune system fights invaders on a daily basis.
These invaders can be fungi, bacteria or viruses.
The innate immune response is its first general 
and immediate strategy, where the acquired immune response
needs time to develop its specialized and more effective
combat forces.

@@

All nucleated cells in humans present randomly sampled polypeptides
fragments to the surroundings of the cell using MHC class I molecules.
If a cell gets infected by a virus, also the virus' polypeptides
will be presented at the cell surface. The foreign viral antigen is detected 
by cytotoxic T lymphocytes, which will kill the infected cell.

@@

All pathogens can be detected by the foreign proteins on 
their (bacterial or fungal) cell walls or (viral) envelope.
For bacteria and fungi, this is the main mode of their detection,
as these do not infect a cell with their (foreign) DNA.
T and B cells express MHC-II to detect foreign polypeptide fragments.
An immune response is started when MHC-II detects a pathogen.

\paragraph{Epitope prediction}

Any human's immune system detects only a fraction of all possible
polypeptide fragments. 
Already each MHC molecule can only bind a
subset of all possible polypeptides.
The classification of HLAs is based on algorithms that
maximize the information content of each 
classification, such as the presence of a certain amino acid at 
the first (for MHC-II, \cite{southwood1998several})
or second position (for MHC-I, \cite{lund2004definition}) of a polypeptide.
However, this does not imply that all polypeptides for a classification
pattern bind to the MHC, as there is no simple method to predict which 
MHCs bind which epitopes.
Using artificial neural networks, it became
possible to make theoretical predictions \cite{nielsen2003reliable,nielsen2004improved},
which are tested to be reliable in practice \cite{larsen2010identification,schellens2008unanticipated,tang2011genome}.

??? 
The polypeptides that can bind are also likelier to by hydrophobic,
as most TMHs have a hydrophobic binding pocket \richel{Reference needed}.
There are two sources of hydrophobic polypeptides, which are 
from membrane proteins and of soluble proteins.
Membrane proteins have at least one TMH to anchor them to
the membrane. TMHs are hydrophobic, as this is required to span the 
hydrophobic cellular lipid membrane. Additionally,
they often have a length of 23 amino acids to be able to span
the membrane.
Soluble proteins have (non-TMH) hydrophobic regions as well, 
to guides the protein to achieve its 3D configuration.

@@

For MHC-I, it was found that predicted epitopes derived 
from human transmembrane helices (TMHs)
are over-presented by all HLA-A and 
most HLA-B super types (\cite{bianchi2017}). 
One explanation is that the presentation of TMHs 
may have an evolutionary advantage for 
the (human) host, as TMHs have a reduced variability 
due to the functional requirement of being able to span a lipid 
bilayer. 
Due to this, pathogens have a lower chance to develop an escape mutation,
as many mutations will result in a disfunctional TMH.
Note that the mechanism by which a cell presents its TMHs is
yet unknown.

@@

If presentation of TMHs on MHC-I would bring an evolutionary advantage 
in the recognition of pathogens by the immune system, 
it would follow that this is equally important for MHC-II, 
especially as the help of CD4+ T cells is needed for a long lasting CD8+ T cell 
response \cite{novy2007cd4}. 
The mechanism to detect foreign TMHs by MHC-II would be unknown, 
similar to the discovery that MHC-I presents TMHs.

@@

The human immune system and human pathogen are in an evolutionary
arms race: our immune systems is selected for the detection
of pathogens, whereas pathogens are selected to avoid detection.
From a pathogen's point of view, however, this struggle 
is of only minor importance:
in seasonal influenza, for example, the selection pressure
exerted by the immune system was only limited \cite{han2019individual}
\richel{but this study does not look into the evolution of TMHs}.
In general, on would expect that evolutionary selection is strongest 
for regions that are important for viral proteins' functions:
in COVID-19, for example, there is preliminary evidence that the strongest
selection pressure is upon residues that changes its virulence \cite{velazquez2020positive}
\richel{residue 251 of ORF3a (an ion channel and a 
viroporin, residue is not part of TMH), and position 84 of ORF8 (unknown function)}.
The part of a pathogen's proteome that is essential, 
only accounts for small part of its full proteome.
Also, these essential parts differ widely between pathogens.
Because of this scarcity and variance in targets, 
one expects that the human immune system
is not tailored to detect these sites, 
as hinted by upon by the aforementioned influenza study.

@@

Regarding human viruses, we do not know the selection pressure
exerted by the human immune system.
For viral TMHs that are undetected by most human haplotypes, we
expect these to be conserved, to remain avoiding detection.
For viral TMHs that are detected by most human haplotypes, we
expect there to be a higher mutation rate, to avoid detection.
Additionally, viruses are not only selected for their evasion of an 
immune response, yet are selected to have a high reproductive value.
Because the strength of the evolutionary signal is unknown,
we use a strong data set to be able to detect it.

@@

Regarding human pathogenic bacteria, we do not know the extent to which the 
human immune systems selects on these, nor which loci are selected
upon \richel{so I should read the literature on that}.
Because most bacterial pathogens are generalists, that is,
can infect multiple hosts, we expect the selection pressure exerted
by the human immune response to be weak.
For bacterial TMHs that are undetected by most human haplotypes, we
expect these to be conserved, to remain avoiding detection.
For bacterial TMHs that are detected by most human haplotypes, we
expect there to be a higher mutation rate, to avoid detection.
Also, it is unknown if bacterial TMHs are detected by MHC-II at all.
Due to this, the strength of the evolutionary signal is unknown,
thus we need a strong data set to be able to detect it.

@@

To detect an evolutionary signal for selection on the TMHs of viruses, 
we use the many time-dated COVID-19 proteomes.
We can assume that only after the moment that COVID-19 spilled over 
to humans, it has mostly been the human immune system selecting against
its detection, although this is only part of the selection for 
having a high transmission rate.
To detect an evolutionary signal for selection on the TMHs of bacteria, 
we use the proteomes of Mycobacterium. Mycobacterium is \richel{something}.

@@

For MHC-II, we do the same analysis as MHC-I, except for
MHC-II haplotypes. We picked the top 21 MHC-II alleles that are most abundant 
in the current human population (\cite{greenbaum2011functional}).
\frans{Show coverage}
We chose to use 21 alleles to have a similar number of alleles as is done
for MHC-I \richel{Unsure if this was the actual reason}.
Same as \cite{bianchi2017}, we defined the 5\% of polypeptides 
with the lowest IC50 values as binders. We then counted the percentage
of binders that have one or more amino acids that are part of a TMH.

@@

We aimed our evolutionary experiment at TMHs, because these can
be predicted well from a protein structure,
are common structures and are present in all pathogens. 
We could have done the same experiment on beta-turn,
as also these can be predicted well \cite{petersen2010netturnp},
are common structures and are present in all pathogens.
